### 2017-10-13 writing/paper/sections/conclfut.tex
Maybe final changes from Henning|

It is interesting to consider the application of the Third Dominating rule, as well as the simple
extension of allowing a sequence of point insertions, to mesh patterns without any dominating
pattern in order to try to capture some of the coincidences described in
\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15} and
\textcite{DBLP:journals/corr/ClaessonTU14}.
 \begin{example}
   We can establish the coincidence between the patterns
   \begin{equation*}
     m_1 = \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1}
   \end{equation*}
   That is not explained by the methods presented by \textcite{DBLP:journals/corr/ClaessonTU14}.

   Consider a permutation containing \(m_1\),
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{}{1,2}{0/1,0/2,1/1,1/2,2/0}
         \draw (1.5,0.5) node {\(Y\)};
         \draw (2.5,1.5) node {\(X\)};
     \end{tikzpicture}
   \end{equation*}
   If the regions corresponding to both \(X\) and \(Y\) are empty then we have
   an occurrence of \(m_2\).
   If the region corresponding to \(X\) is non-empty, we can then choose
   the lowest valued point in this region
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{1,3}{1,3,2}{0/1,0/2,0/3,1/1,1/2,1/3,2/0,2/1,3/0,3/1}
         \draw (1.5,0.5) node {\(Y\)};
     \end{tikzpicture}
   \end{equation*}
   If the region corresponding to \(Y\) is empty then we have an occurrence of
   \(m_2\) with the indicated points.
   Now if the region corresponding to \(Y\) is non-empty, we can choose the
   rightmost point in this region.
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{2,4}{2,1,4,3}{0/2,0/3,0/4,1/2,1/3,1/4,2/0,2/1,2/2,2/3,2/4,3/0,3/1,3/2,4/0,4/1,4/2}
     \end{tikzpicture}
   \end{equation*}
   And now the two indicated points form an occurrence of \(m_2\).
   We have therefore shown that any occurrence of \(m_1\) leads to an occurrence of
   \(m_2\) and we can easily show the converse by the same reasoning, so \(m_1\)
   and \(m_2\) are coincident.
   This is captured by an extension of the Third Dominating rule where we allow
   multiple steps of adding points before we check for subpattern containment.
 \end{example}

>>>>

It is interesting to consider the application of the Third Dominating rule, as well as the simple
extension of allowing a sequence of point insertions, to mesh patterns without any dominating
pattern in order to try to capture some of the coincidences described in
\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15} and
\textcite{DBLP:journals/corr/ClaessonTU14}.
 \begin{example}
   The coincidence of the patterns
   \begin{equation*}
     m_1 = \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1}
   \end{equation*}
	does not follow from the general methods presented by \textcite{DBLP:journals/corr/ClaessonTU14}, but is rather handled there as
	a special case. We can do it as follows:
   Consider a permutation containing \(m_1\),
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{}{1,2}{0/1,0/2,1/1,1/2,2/0}
         \draw (1.5,0.5) node {\(Y\)};
         \draw (2.5,1.5) node {\(X\)};
     \end{tikzpicture}
   \end{equation*}
   If the regions corresponding to both \(X\) and \(Y\) are empty then we have
   an occurrence of \(m_2\).
   If the region corresponding to \(X\) is non-empty, we can then choose
   the lowest valued point in this region
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{1,3}{1,3,2}{0/1,0/2,0/3,1/1,1/2,1/3,2/0,2/1,3/0,3/1}
         \draw (1.5,0.5) node {\(Y\)};
     \end{tikzpicture}
   \end{equation*}
   If the region corresponding to \(Y\) is empty then we have an occurrence of
   \(m_2\) with the indicated points.
   Now if the region corresponding to \(Y\) is non-empty, we can choose the
   rightmost point in this region.
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{2,4}{2,1,4,3}{0/2,0/3,0/4,1/2,1/3,1/4,2/0,2/1,2/2,2/3,2/4,3/0,3/1,3/2,4/0,4/1,4/2}
     \end{tikzpicture}
   \end{equation*}
   And now the two indicated points form an occurrence of \(m_2\).
   We have therefore shown that any occurrence of \(m_1\) leads to an occurrence of
   \(m_2\) and we can easily show the converse by the same reasoning, so \(m_1\)
   and \(m_2\) are coincident.
   This is captured by an extension of the Third Dominating rule where we allow
   multiple steps of adding points before we check for subpattern containment.
 \end{example}

### 2017-10-13 writing/paper/sections/wilf.tex
Maybe final changes from Henning|

Wilf-equivalence is an important aspect to study in the field of permutation
patterns.

@@

Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
must only use symmetries that preserve the dominating pattern(s) in \(S\), if we
were to allow other symmetries, then the equivalences calculated in the
previous section do not necessarily hold.

Throughout this section we will consider Wilf-equivalences of patterns
whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\), if two patterns \(p_1 \et p_2\) are Wilf-equivalent we will denote this \(p_1\weqv{231}p_2\), similarly when computational methods indicate that two patterns are Wilf equivalent under \(231\) we write \(p_1\weqv[comp]{231}p_2\). We will
use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
be the usual Catalan generating function satisfying \(C(x) = 1 + xC(x)^2\).
The fact that \(C(x)\) is the generating function for \(\mathcal{C}\)
can be seen by structural decomposition around the maximum, as
shown in \FigureRef{fig:decompmax}.

@@

\subsection{Wilf-classes with mesh patterns of length 1.}
\label{wilf1}
When considering the mesh patterns of length \(2\) it will be useful to know
the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
\(\av{\perm{2,3,1}}\), this means that we are considering the set
\(\av{\{\perm{2,3,1},p}\}\)
where \(p\) is a mesh-pattern of length 1.

@@

Computing avoiders up to length 10 suggests that there are at least 23 \(\weqv{231}\)-equivalence classes,
of which 13 are not equal to coincidence classes (after taking symmetries). Therefore there seem to be Wilf-equivalences that are not
explained by coincidences or symmetry.

@@

If there is an occurrence of \(m_1\) there must be a lowest occurrence of \(m_1\), \ie, the occurrence has the lowest possible values for any occurrence. Now consider the top
regions, labelled \(A \et B\), the subpermutation contained in these regions must avoid the permutation \(231\). Also, the subpermutation in the region \(A\) must be a decreasing subsequence, otherwise an occurrence of \(231\) will be created with either of the points in the occurrence. Now, consider the region labelled \(C\), since we specified that we were focused on the lowest possible occurrence of \(m_1\) this region cannot contain an occurrence of either \(m_1\) or \(231\). We now have a structural decomposition of a container of \(m_1\) inside \(\av{231}\).

@@

Similarly to a container of \(m_1\) we will look at the lowest possible occurrence of \(m_2\), similar to before the regions \(A^\prime \et B^\prime\) together contain an avoider of \(231\) and \(A\) contains a decreasing subsequence. Since we are considering the lowest occurrence of \(m_2\) so the region \(C^\prime\) does not contain an occurrence of either \(m_2\) or \(231\).

@@

Consider the class, \(\scriptvar_3\), of permutations defined by avoiding \(\perm{2,3,1}\) and \(m_3\). We can decompose a member of this class around the maximum
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_1}
\end{equation*}
Only the first point in the top right region can
create an occurrence of \(m_3\) if and only if it is the
element with largest value in this region, therefore the partial permutation
in this region must avoid starting with the maximum.

Looking at avoiders of \(\perm{2,3,1}\) and \(m_4\) we
can perform a similar decomposition around the maximum to get
\begin{equation*}
    \scriptvar_4 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\scriptvar_4}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
An occurrence of \(m_4\) can never occur in the
top right region. It could only occur between the maximum and the first point
in the region, if and only if this first point is the lowest valued element in
this region, so this top right region must contain a sub-permutation that does not
start with it's minimum, i.e., it is a member of \(\mathcal{\varmaxl}_3\) described
in \SectionRef{wilf1}.
Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same
enumeration, \(\scriptvar_3\) and \(\scriptvar_4\) must also have
the same enumeration and are therefore \(m_3\weqv{m_4}\).

Now we must consolidate these two subclasses. In order to do this we must
consider the decomposition around the leftmost point of a permutation in
\(\av{\{\perm{2,3,1},m_1\}}\). We have the following
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompleft{1/0,0/1}{\scriptvar_1}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
have the same enumeration, and therefore \(m_1 \weqv{231} m_2 \weqv{231}m_3\weqv{m_4}\) with generating function satisfying
\begin{equation*}
    \var(x) = 1 + x\var(x)\varmaxl(x)
\end{equation*}
where \(\varmaxl(x)\) is the generating function given in \EquationRef{eqn:maxlgen}.
This can be enumerated to give the sequence
\begin{equation*}
    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc \tag{\href{https://oeis.org/A035929}{OEIS: A035929} offset 1}
\end{equation*}

@@

In order to establish the remainder of the Wilf-equivalences of the form \(\av{\{\perm{2,3,1},p\}}\) where \(p\)
is a mesh pattern we can use similar methods to allow us to consolidate experimental classes into
actual classes, these methods allow us to explain all 23 of the observed Wilf-classes seen in experimentation.

>>>>

Wilf-equivalence is an important aspect to study in the field of permutation
patterns. The original definition is as follows:

@@

Since we will consider Wilf-equivalences in a set \(\av{S}\) we
must only use symmetries that preserve the dominating pattern(s) in \(S\), if we
were to allow other symmetries, then the equivalences calculated in the
previous section do not necessarily hold.

In the remainder of the paper we will consider Wilf-equivalences of patterns
whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\), if two patterns \(p_1 \et p_2\) are Wilf-equivalent we will denote this \(p_1\weqv{231}p_2\), similarly when computational methods indicate that two patterns are Wilf equivalent under \(231\) we write \(p_1\weqv[comp]{231}p_2\). We will
use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
be the usual Catalan generating function satisfying \(C(x) = 1 + xC(x)^2\).
The fact that \(C(x)\) is the generating function for \(\mathcal{C}\)
can be seen by structural decomposition around the maximum, as
shown in \FigureRef{fig:decompmax}.

@@

\subsection{Wilf-classes with mesh patterns of length 1.}
\label{wilf1}
When considering the mesh patterns of length \(2\) it will be useful to know
the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
\(\av{\perm{2,3,1}}\), this means that we are considering the \(\weqv{231}\)-equivalence classes of mesh patterns with underlying classical pattern \(1\).

@@

Computing avoiders up to length 10 gives 23 \(\weqv[comp]{231}\)-equivalence classes,
of which 13 are not equal to coincidence classes (after taking symmetries). Therefore there seem to be Wilf-equivalences that are not
explained by coincidences or symmetry.

@@

If there is an occurrence of \(m_1\) there must be a lowest occurrence of \(m_1\), \ie, the occurrence has the lowest possible values for any occurrence. Now consider the top
regions, labelled \(A \et B\), the subpermutation contained in the union of these regions must avoid the permutation \(231\). Also, the subpermutation in the region \(A\) must be a decreasing subsequence, otherwise an occurrence of \(231\) will be created with either of the points in the occurrence of \(m_1\). Now, consider the region labelled \(C\), since we specified that we were focused on the lowest possible occurrence of \(m_1\) this region cannot contain an occurrence of either \(m_1\) or \(231\). This is a full structural decomposition of a container of \(m_1\) inside \(\av{231}\).

@@

Similarly to a container of \(m_1\) we will look at the lowest possible occurrence of \(m_2\), and as before the regions \(A^\prime \et B^\prime\) together contain an avoider of \(231\) and \(A\) contains a decreasing subsequence. Since we are considering the lowest occurrence of \(m_2\) the region \(C^\prime\) does not contain an occurrence of either \(m_2\) or \(231\).

@@

Consider the class, \(\scriptvar_3\), of permutations defined by avoiding \(\perm{2,3,1}\) and \(m_3\). We can decompose a member of this class around the maximum
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_1}
\end{equation*}
Only the first point in the top right region can
create an occurrence of \(m_3\) if and only if it is the
element with largest value in this region, therefore the partial permutation
in this region must avoid starting with the maximum, \ie, be in \(\mathcal{\varmaxl}_1 = \av{\textpattern{}{1}{0/0,0/1,1/1}}\) described
in \SectionRef{wilf1}.

Looking at avoiders of \(\perm{2,3,1}\) and \(m_4\) we
can perform a similar decomposition around the maximum to get
\begin{equation*}
    \scriptvar_4 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\scriptvar_4}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
An occurrence of \(m_4\) can never occur in the
top right region. It could only occur between the maximum and the first point
in the region, if and only if this first point is the lowest valued element in
this region, so this top right region must contain a sub-permutation that does not
start with it's minimum, i.e., it is a member of \(\mathcal{\varmaxl}_3 = \av{\av{\textpattern{}{1}{0/0,0/1,1/0}}}\).
Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same
enumeration, \(\scriptvar_3\) and \(\scriptvar_4\) must also have
the same enumeration and are therefore \(m_3\weqv{231}m_4\).

Now we must consolidate these two subclasses. In order to do this we must
consider the decomposition around the leftmost point of a permutation in
\(\av{\{\perm{2,3,1},m_1\}}\). We have the following
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompleft{1/0,0/1}{\scriptvar_1}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
It is therefore clear that avoiders of \(m_1\) and avoiders of \(m_4\)
have the same enumeration, and therefore \(m_1 \weqv{231} m_2 \weqv{231}m_3\weqv{m_4}\) with generating function satisfying
\begin{equation*}
    \var(x) = 1 + x\var(x)\varmaxl(x)
\end{equation*}
where \(\varmaxl(x)\) is the generating function given in \EquationRef{eqn:maxlgen}.
This can be enumerated to give the sequence
\begin{equation*}
    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc \tag{\href{https://oeis.org/A035929}{OEIS: A035929} offset 1}
\end{equation*}

@@

In order to establish the remainder of the Wilf-equivalences of the form \(\av{\{\perm{2,3,1},p\}}\) where \(p\)
is a mesh pattern of length \(2\) we can use similar methods to allow us to consolidate experimental classes into
actual classes, these methods allow us to explain all 23 of the observed Wilf-classes seen in experimentation.
